---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

Getting the required packages for data cleaning and exploration

```{r}
library(tidyverse)
library(ggplot2)
library(skimr)
```

## Cleaning the data

First, the datasets "train_values" and train_labels" are imported


```{r}
library(janitor)
library(here)
```

Taking closer looks into the data:

```{r}
install.packages("skimr")
library(skimr)
```

# Looking for missing values:

```{r}
train_values %>% 
  filter(!complete.cases(.))
```

## Changing the variable from 'chr' to 'factor' affecting column land_surface_condition

```{r}
#confirming the datatype of the column 'land_surface_condition'
class(train_values$land_surface_condition)

#categorizing the values in the column
unique(train_values$land_surface_condition)

#converting the character type to a categorical type: factor
train_values$land_surface_condition <- as.factor(train_values$land_surface_condition)
class(train_values$land_surface_condition)
levels(train_values$land_surface_condition)
```

## Changing the variable from 'chr' to 'factor' affecting columns from 9 to 15

```{r}
#for foundation_type
unique(train_values$foundation_type)
train_values$foundation_type <- as.factor(train_values$foundation_type)
class(train_values$foundation_type)
levels(train_values$foundation_type)

#for roof_type
unique(train_values$roof_type)
train_values$roof_type <- as.factor(train_values$roof_type)
class(train_values$roof_type)
levels(train_values$roof_type)

#for ground_floor_type
unique(train_values$ground_floor_type)
train_values$ground_floor_type <- as.factor(train_values$ground_floor_type)
class(train_values$ground_floor_type)
levels(train_values$ground_floor_type)

#for other_floor_type
unique(train_values$other_floor_type)
train_values$other_floor_type <- as.factor(train_values$other_floor_type)
class(train_values$other_floor_type)
levels(train_values$other_floor_type)

#for position
unique(train_values$position)
train_values$position <- as.factor(train_values$position)
class(train_values$position)
levels(train_values$position)

#for plan_configuration
unique(train_values$plan_configuration)
train_values$plan_configuration <- as.factor(train_values$plan_configuration)
class(train_values$plan_configuration)
levels(train_values$plan_configuration)
```

## Grouping the building_ids with oter parameters:

Grouping the building_ids for each unique count_floors_pre_eq):
```{r}
floor_count <- train_values %>% 
  select(building_id,count_floors_pre_eq) %>% 
  group_by(count_floors_pre_eq) %>%
  summarise(building_id = n())
View(floor_count)
```

Grouping the building_ids for each unique age):
```{r}
age_count <- train_values %>% 
  select(building_id,age) %>% 
  group_by(age) %>%
  summarise(building_id = n())
View(age_count)
```

Grouping the building_ids for each unique land_surface_condition):
```{r}
land_count <- train_values %>% 
  select(building_id,land_surface_condition) %>% 
  group_by(land_surface_condition) %>%
  summarise(building_id = n())
View(land_count)
```

Grouping the building_ids for each unique plan_configuration):
```{r}
plan_count <- train_values %>% 
  select(building_id,plan_configuration) %>% 
  group_by(plan_configuration) %>%
  summarise(building_id = n())
View(plan_count)
```


## Joining the datasets: train_values and train_labels for analysis:

```{r}
joined_train_dataset <- inner_join(train_values, train_labels, by = c("building_id"))
View(joined_train_dataset)
```


```{r}
library(ggplot2)
ggplot(joined_train_dataset, aes(x = factor(damage_grade))) +
  geom_bar(color = "white") +
  labs(title = "Histogram of Damage Grades",
       x = "Damage Grade",
       y = "Number of Buildings") 
```



```{r}
library(GGally)
```

# Scatter plot of age vs. damage_grade
```{r}
ggplot(joined_train_dataset, aes(x = age, y = damage_grade, color = factor(damage_grade))) +
  geom_point(size = 5, alpha = 0.7) +
  labs(title = "Scatter Plot of Age vs. Damage Grade",
       x = "Age",
       y = "Damage Grade") +
  scale_color_manual(values = colors)
```

# Create a subset of the joined dataset with relevant columns
```{r}
subset <- joined_train_dataset %>%
  select(c("damage_grade", "age", "area_percentage", "height_percentage"))
```

# Histogram for age with color-coded damage_grade
```{r}
ggplot(subset, aes(x = age, fill = factor(damage_grade))) +
  geom_histogram(binwidth = 15, position = "identity", alpha = 1) +
  labs(title = "Histogram for Age with Color-Coded Damage Grade",
       x = "Age",
       y = "Frequency",
       fill = "Damage Grade") +
  scale_fill_manual(values = colors)
```



# Scatter plot for area_percentage vs. height_percentage with color-coded damage_grade
```{r}
ggplot(subset, aes(x = area_percentage, y = height_percentage, color = factor(damage_grade))) +
  geom_point(alpha = 0.7, size = 3) +
  labs(title = "Scatter Plot for Area Percentage vs. Height Percentage with Color-Coded Damage Grade",
       x = "Area Percentage",
       y = "Height Percentage",
       color = "Damage Grade") +
  scale_color_manual(values = colors)
```


## Scatter plot matrix for age,area_percentage, height_percentage, count_floors and damage_grade
```{r}
pairs(joined_train_dataset[, c("damage_grade", "count_floors_pre_eq", "age", "area_percentage", "height_percentage")], 
      col = factor(joined_train_dataset$damage_grade))
```



# Create a subset of the joined dataset with relevant columns
```{r}
subset_data <- joined_train_dataset %>%
  select(c("damage_grade", "has_superstructure_rc_non_engineered", "has_superstructure_rc_engineered","foundation_type", 
                     "area_percentage", 
                     "height_percentage",
                     "count_floors_pre_eq",
                     "land_surface_condition",
                     "has_superstructure_cement_mortar_stone"))
```


# Pair plot with different colors for damage_grade
```{r}
pairs(subset_data[, c("area_percentage", 
                     "height_percentage", "has_superstructure_rc_engineered", 
                     "count_floors_pre_eq",
                     "land_surface_condition",
                     "has_superstructure_cement_mortar_stone")], 
      col = factor(subset_data$damage_grade), 
      pch = 16)
```


